---
title: "Supreme Court modelling"
author: "Austin Spivey"
date: "February 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Supreme Court Modelling

<br>

To-Do:

* Load packages
* Load in raw data
* Run descriptive stats and analyze summary
* Clean and wrangle, as needed
* First attempts at modelling



<br>



### Load packages

<br>

```{r}
# install.packages("tidyverse")
# install.packages("caret")
# install.packages("caTools")
# install.packages("rpart")
# install.packages("rpart.plot")
# install.packages("randomForest")
# install.packages("e1071")
# install.packages("scales")


library(tidyverse)
library(caret)
library(caTools)
library(rpart)
library(rpart.plot)
library(randomForest)
library(e1071)
library(scales)
```



<br>



### Load in raw data

<br>

```{r}
folder_path = "C:\\Users\\aspivey\\Desktop\\Spivey Stuff\\Supreme Court Data\\SCDB_2018_02_justiceCentered_LegalProvision.csv\\SCDB_2018_02_justiceCentered_LegalProvision.csv"

raw_df = read.csv2(folder_path, header = TRUE, sep = ",", stringsAsFactors = FALSE)
```



<br>



### Run descriptive stats and analyze summary

<br>

```{r}
str(raw_df)
```


```{r}
raw_df$dateArgument = as.Date(raw_df$dateArgument, "%m/%d/%Y")
raw_df$dateRearg = as.Date(raw_df$dateRearg, "%m/%d/%Y")
raw_df$dateDecision = as.Date(raw_df$dateDecision, "%m/%d/%Y")
```

<br>






Examine Chiefs:

```{r}
# How many SCOTUS Chief Justices have there been since 1946?
# What was the average number of years spent as Chief Justice?
# Who had the longest stint as Chief? The shortest? 

data.frame(distinct((raw_df %>% group_by(chief)), chief, "Years on Bench" = paste0(min(term),"-",max(term)), "Length of Term" = max(term)-min(term)), "Average Years on Bench for All" = colMeans(distinct((raw_df %>% group_by(chief)), max(term)-min(term))[2]))


raw_df %>% group_by(justice, direction) %>% select(justice, justiceName, direction) %>% mutate(count = n()) %>% distinct() %>% arrange(justice) %>% group_by(justice) %>%
  mutate(count_all = sum(count), percentage = percent(count / count_all))


# Whose votes have been most "conservative"? Most "liberal"?

raw_df %>%
  group_by(justice, direction) %>%
  select(justice, justiceName, direction) %>%
  mutate(count = n()) %>%
  distinct() %>%
  arrange(justice) %>%
  group_by(justice) %>%
  mutate(count_all = sum(count), percentage = percent(count / count_all)) %>%
  ungroup() %>%
  group_by(direction) %>%
  filter(percentage == max(percentage))
```






Examine issues:

```{r}
# How many unique dockets have there been since 1946?

raw_df %>% group_by(docketId) %>% select(docketId) %>% count() %>% arrange(desc(n))

# How many unanimous decisions have there been?

raw_df %>% group_by(caseId) %>% select(caseId, caseIssuesId, majVotes, minVotes) %>% distinct() %>% filter(majVotes == 0 | minVotes == 0) %>% arrange(majVotes)

# What date had the greatest number of oral arguments?

raw_df %>%
  group_by(dateArgument) %>%
  select(dateArgument, caseId) %>%
  distinct() %>%
  mutate(count = n()) %>%
  select(-caseId) %>%
  distinct() %>%
  arrange(desc(count)) %>%
  ungroup() %>%
  filter(is.na(dateArgument) == F) %>%
  filter(count == max(count))




raw_df %>%
  select(dateArgument, caseId, docketId, caseIssuesId) %>%
  filter(dateArgument == "1947-10-13") %>%
  distinct()
```










### Clean and wrangle, as needed

<br>

```{r}
fewvar_df = raw_df
```


```{r}
# # Remove:
# 
#   # ID vars:
#       caseId
#       docketId
#       caseIssuesId
#       voteId
#       usCite
#       sctCite
#       ledCite
#       lexisCite
#       docket
# 
#   # Outcome vars:
#       decisionType
#       declarationUncon
#       caseDisposition
#       caseDispositionUnusual
#       precedentAlteration
# 
#   # Voting & Opinion vars:
#       voteUnclear
#       majOpinWriter
#       majOpinAssigner
#       splitVote
#       majVotes
#       minVotes
#       vote
#       opinion
#       direction
#       majority
#       firstAgreement
#       secondAgreement

```

<br>

```{r}
fewvar_df = subset(raw_df, select=-c(caseId,
                       docketId,
                       caseIssuesId,
                       voteId,
                       usCite,
                       sctCite,
                       ledCite,
                       lexisCite,
                       docket,
                       decisionType,
                       declarationUncon,
                       caseDisposition,
                       caseDispositionUnusual,
                       precedentAlteration,
                       voteUnclear,
                       majOpinWriter,
                       majOpinAssigner,
                       splitVote,
                       majVotes,
                       minVotes,
                       vote,
                       opinion,
                       direction,
                       majority,
                       firstAgreement,
                       secondAgreement
                       ))
```




<br>






### First attempts at modelling

<br>


Split dataset

```{r}
set.seed(999)
split =sample.split(fewvar_df$partyWinning, SplitRatio =0.7)
train =subset(fewvar_df, split ==TRUE)
test =subset(fewvar_df, split ==FALSE)
```

<br>


For this model, I want to predict on the variable `partyWinning`, which indicates whether the petitioning party (i.e., the plaintiff or the appellant) emerged victorious.  

The possible values:

* **0** -	no favorable disposition for petitioning party apparent
* **1** - 	petitioning party received a favorable disposition
* **2** - 	favorable disposition for petitioning party unclear


<br>


Create a classification tree to predict `partyWinning` (indicates whether the petitioning party (i.e., the plaintiff or the appellant) emerged victorious). Use the training set and all variables. Plot the tree

```{r}
fit = rpart(partyWinning ~ ., data = train, method = "class")
prp(fit)
```

